---
title: "proj2"
author: "Kwong Yu Chong, Jiahua Wang"
date: "2022-11-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
require(tidyverse)
require(kableExtra)
require(patchwork)
require(GGally)
require(corrplot)
```

# Read Data
```{r}
img1 = read.table("imagem1.txt")
img2 = read.table("imagem2.txt")
img3 = read.table("imagem3.txt")
```

```{r}
col_name = c("Y","X","Label","NDAI","SD","CORR","Rad_Df",
             "Rad_Cf","Rad_Bf","Rad_Af","Rad_An")
colnames(img1) = col_name
colnames(img2) = col_name
colnames(img3) = col_name

img1$image = 1
img2$image = 2
img3$image = 3

full_img = rbind(img1, img2, img3)
```

```{r}
full_img %>% head()
```

# Summarize the data

###  % pixcel by label
```{r}
tbl1= full_img %>% filter(image==1) %>% group_by(Label) %>% count() %>%
  ungroup() %>% 
  mutate(Image1 = round(n / sum(n)*100,1)) %>% select(Label, Image1)
tbl2= full_img %>% filter(image==2) %>% group_by(Label) %>% count() %>%
  ungroup() %>% 
  mutate(Image2 = round(n / sum(n)*100,1)) %>% select(Label, Image2)
tbl3= full_img %>% filter(image==3) %>% group_by(Label) %>% count() %>%
  ungroup() %>% 
  mutate(Image3 = round(n / sum(n)*100,1)) %>% select(Label, Image3)

tbl1 %>% left_join(tbl2, by='Label') %>% left_join(tbl3, by='Label') %>%
  mutate(Label=ifelse(Label==1, 'Cloudy (1)', ifelse(Label==-1, 'Cloud-free (-1)', 'Unlabeled (0)')))%>%
  arrange(Label)%>%
  kbl(caption = "<center><strong>Pixels By Label (%)<center><strong>",
      escape = FALSE,
      format = 'html',) %>%
  kable_classic(full_width = F, html_font = "Cambria")
```

### Plot Maps
```{r}
full_img %>% filter(image==1) %>% 
  mutate(Label=
           as.factor(ifelse(Label==1, 'Cloudy (1)',
                  ifelse(Label==-1, 'Cloud-free (-1)',
                         'Unlabeled (0)')))) %>% 
  ggplot() +
  geom_raster(aes(x=X, y=Y, fill=Label)) +
  scale_fill_manual(values=c("Cloudy (1)"='white', 
                             'Cloud-free (-1)'='grey',
                             'Unlabeled (0)'='black'))+
  labs(title='Image 1 Label', x='X-coordinate', y='Y-coordinate')+
   theme(plot.title = element_text(hjust = 0.5), legend.position="none")

full_img %>% filter(image==2) %>% 
  mutate(Label=
           as.factor(ifelse(Label==1, 'Cloudy (1)',
                  ifelse(Label==-1, 'Cloud-free (-1)',
                         'Unlabeled (0)')))) %>% 
  ggplot() +
  geom_raster(aes(x=X, y=Y, fill=Label)) +
  scale_fill_manual(values=c("Cloudy (1)"='white', 
                             'Cloud-free (-1)'='grey',
                             'Unlabeled (0)'='black'))+
  labs(title='Image 2 Label', x='X-coordinate', y='Y-coordinate')+
   theme(plot.title = element_text(hjust = 0.5), legend.position="none")

full_img %>% filter(image==3) %>% 
  mutate(Label=
           as.factor(ifelse(Label==1, 'Cloudy (1)',
                  ifelse(Label==-1, 'Cloud-free (-1)',
                         'Unlabeled (0)')))) %>% 
  ggplot() +
  geom_raster(aes(x=X, y=Y, fill=Label)) +
  scale_fill_manual(values=c("Cloudy (1)"='white', 
                             'Cloud-free (-1)'='grey',
                             'Unlabeled (0)'='black'))+
  labs(title='Image 3 Label', x='X-coordinate', y='Y-coordinate')+
   theme(plot.title = element_text(hjust = 0.5),legend.position="none")
```

# EDA
```{r}
col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
full_img %>% select(-c(X,Y,Label, image)) %>% cor() %>%
  corrplot::corrplot(method="color", col=col(200),  
     diag=T,  tl.srt=45, tl.col = 'black', tl.cex = 0.75,mar=c(0,0,2,0),
     type="upper", order="hclust", 
     title='Feature Correlation', 
         addCoef.col = "black",  outline=T,
         )
```

```{r}
tbl_data = full_img %>% select(-image)%>% filter(Label!=0) %>% cor() %>% .[-c(1,2,3),3] %>% round(.,2)

data.frame('value'=tbl_data) %>% arrange(desc(value))%>%kbl(
  caption = "<center><strong>Correlation with Label<center><strong>",
  escape = FALSE,
  format = 'html',
  table.attr = "style='width:25%;'") %>%
  kable_classic(full_width = T, html_font = "Cambria",) 
```

```{r}
plot_data = full_img %>% 
  filter(Label != 0) %>%
  mutate(Label = ifelse(Label==1, 'Cloudy', 'Cloud-free'))

p1 = plot_data%>%
  ggplot() + geom_density(aes(x=CORR, fill=Label), alpha=0.5) 
p2 = plot_data%>%
  ggplot() + geom_density(aes(x=NDAI, fill=Label), alpha=0.5) 
p3 = plot_data%>%
  ggplot() + geom_density(aes(x=log(SD), fill=Label), alpha=0.5) 
p4 = plot_data%>%
  ggplot() + geom_density(aes(x=Rad_Af, fill=Label), alpha=0.5) 

p1 + p2 + p3 + p4 + plot_layout(nrow=2, guides = "collect") + 
  plot_annotation(title = 'Density of Features By Label',
                  theme=theme(plot.title = element_text(hjust = 0.5))) 
```

## Two-sample Kolmogorov–Smirnov statistics
```{r}
full_img_cloudy = full_img %>% select(-image)%>%filter(Label==1)
full_img_not_cloudy = full_img %>% select(-image)%>% filter(Label==-1)

var_names = full_img %>% select(-c(X, Y, Label, image)) %>% colnames() 
KS_stat = matrix(nrow=length(var_names))
rownames(KS_stat) = var_names
colnames(KS_stat) = 'Statistics'

for (var in var_names){
  cloudy_density = full_img_cloudy %>% select(var) %>% as.matrix()%>% density()
  not_cloudy_density = full_img_not_cloudy %>% select(var)%>% as.matrix()%>% density()
  min_x = min(cloudy_density$x , not_cloudy_density$x)
  max_x = max(cloudy_density$x , not_cloudy_density$x)
  cloudy_density = full_img_cloudy %>% select(var) %>% as.matrix()%>% 
    density(from=min_x, to=max_x)
  not_cloudy_density = full_img_not_cloudy %>% select(var)%>% as.matrix()%>%
    density(from=min_x, to=max_x)
  
  x_diff = not_cloudy_density$x %>% diff() %>% .[1]
  KS = tibble(x=cloudy_density$x, cloudy=cloudy_density$y, not_cloudy= not_cloudy_density$y)%>%
    mutate(cloudy = cumsum(cloudy *x_diff),
           not_cloudy= cumsum(not_cloudy *x_diff))%>%
    mutate(diff = abs(cloudy - not_cloudy))%>%
    select(diff)%>%max()
  KS_stat[var,1] = KS
}
KS_stat %>% round(.,2) %>% data.frame() %>% arrange(desc(Statistics))%>%kbl(
  caption = "<center><strong>Two-sample K-S Test: cloudy VS cloud-free<center><strong>",
  escape = FALSE,
  format = 'html',
  table.attr = "style='width:45%;'") %>%
  kable_classic(full_width = T, html_font = "Cambria",) 
```

## PCA
Removed unlabeled data
```{r}
pca_data = full_img %>%select(-c(X,Y,Label, image))
pca = prcomp(pca_data, scale.  =T)
explained_var = (pca$sdev^2) %>% cumsum()
explained_var = round(explained_var / explained_var[8]*100,1)

ggplot()+
  geom_point(aes(x=seq(1,8), y=pca$sdev^2))+
  geom_line(aes(x=seq(1,8), y=pca$sdev^2))+
  labs(x='PC', y='Variance', title='Screeplot')
```

*PC1 is mostly radiance variables. Look at principal direction.
```{r}
full_pca = tibble(data.frame(pca$x))
full_pca$X = full_img$X
full_pca$Y = full_img$Y
full_pca$Label = ifelse(full_img$Label==1,'Cloudy',ifelse(full_img$Label==-1,'Cloud-free',
                                                               'Unlabeled'))
full_pca$image = full_img$image


p1 = full_pca%>% filter(Label!='Unlabeled')%>%ggplot()+
  geom_density(aes(x=PC1, fill=Label), alpha=0.3)+
  labs(title=paste0('PC1(',explained_var[1],'%)'), y='Density', x='')+
  theme(plot.title = element_text(hjust = 0.5))



p2 = full_pca%>% filter(Label!='Unlabeled')%>%ggplot()+
  geom_density(aes(x=PC2, fill=Label), alpha=0.3)+
  labs(title =paste0('PC2 (',explained_var[2]-explained_var[1],'%)'), y='Density', x='')+
  theme(plot.title = element_text(hjust = 0.5))


p3=full_pca %>% filter(image==1) %>% 
  ggplot() +
  geom_raster(aes(x=X, y=Y, fill=PC1)) +
  #labs(title='Image 1: PC1', x='X-coordinate', y='Y-coordinate')+
   theme(plot.title = element_text(hjust = 0.5), legend.position="none")

p4=full_pca %>% filter(image==1) %>% 
  ggplot() +
  geom_raster(aes(x=X, y=Y, fill=PC2)) +
  #labs(title='Image 1: PC2', x='X-coordinate', y='Y-coordinate')+
   theme(plot.title = element_text(hjust = 0.5), legend.position="none")




p1 + p2 + p3+p4+plot_layout(nrow=2, guides = "collect")
```
```{r}
p1=full_pca %>% filter(image==1) %>% 
  ggplot() +
  geom_raster(aes(x=X, y=Y, fill=PC2)) +
  labs(title='Image 1: PC1', x='', y='')+
   theme(plot.title = element_text(hjust = 0.5), legend.position="none",
         axis.text        = element_blank(),
        axis.ticks       = element_blank(),
        axis.title       = element_blank(),
        panel.background = element_blank())


p2=full_pca %>% filter(image==1) %>% 
  ggplot() +
  geom_raster(aes(x=X, y=Y, fill=PC2)) +
  labs(title='Image 1: PC2', x='', y='')+
   theme(plot.title = element_text(hjust = 0.5), legend.position="none",
         axis.text        = element_blank(),
        axis.ticks       = element_blank(),
        axis.title       = element_blank(),
        panel.background = element_blank())
p3 = full_img %>% filter(image==1) %>% 
  ggplot() +
  geom_raster(aes(x=X, y=Y, fill=NDAI)) +
  labs(title='Image 1: NDAI', x='', y='')+
   theme(plot.title = element_text(hjust = 0.5), legend.position="none",
         axis.text        = element_blank(),
        axis.ticks       = element_blank(),
        axis.title       = element_blank(),
        panel.background = element_blank())
p4 = full_img %>% filter(image==1) %>% 
  ggplot() +
  geom_raster(aes(x=X, y=Y, fill=SD)) +
  labs(title='Image 1: SD', x='', y='')+
   theme(plot.title = element_text(hjust = 0.5), legend.position="none",
         axis.text        = element_blank(),
        axis.ticks       = element_blank(),
        axis.title       = element_blank(),
        panel.background = element_blank())

p1+p2+p3+p4 + plot_layout(nrow=2)
```

```{r}
full_pca %>% select(-image)%>% filter(Label!="Unlabeled")%>% 
  mutate(Label=ifelse(Label=='Cloudy', 1, -1)) %>%
  cor() %>% .[1:8,11] %>% round(.,2) %>% 
  data.frame('value'=.) %>% arrange(desc(value))%>%kbl(
  caption = "<center><strong>Correlation with Label<center><strong>",
  escape = FALSE,
  format = 'html',
  table.attr = "style='width:25%;'") %>%
  kable_classic(full_width = T, html_font = "Cambria",) 
```

```{r}
full_PC = data.frame(pca$x) %>% as.tibble() %>% mutate(Label=full_img$Label)
full_PC_cloudy = full_PC %>% filter(Label==1)
full_PC_not_cloudy = full_PC %>% filter(Label==-1)

var_names = full_PC %>% select(-Label)%>% colnames() 
KS_stat = matrix(nrow=length(var_names))
rownames(KS_stat) = var_names
colnames(KS_stat) = 'Statistics'

for (var in var_names){
  cloudy_density = full_PC_cloudy %>% select(var) %>% as.matrix()%>% density()
  not_cloudy_density = full_PC_not_cloudy %>% select(var)%>% as.matrix()%>% density()
  min_x = min(cloudy_density$x , not_cloudy_density$x)
  max_x = max(cloudy_density$x , not_cloudy_density$x)
  cloudy_density = full_PC_cloudy %>% select(var) %>% as.matrix()%>% 
    density(from=min_x, to=max_x)
  not_cloudy_density = full_PC_not_cloudy %>% select(var)%>% as.matrix()%>%
    density(from=min_x, to=max_x)
  
  x_diff = not_cloudy_density$x %>% diff() %>% .[1]
  KS = tibble(x=cloudy_density$x, cloudy=cloudy_density$y, not_cloudy= not_cloudy_density$y)%>%
    mutate(cloudy = cumsum(cloudy *x_diff),
           not_cloudy= cumsum(not_cloudy *x_diff))%>%
    mutate(diff = abs(cloudy - not_cloudy))%>%
    select(diff)%>%max()
  KS_stat[var,1] = KS
}
KS_stat %>% round(.,2) %>% data.frame() %>% arrange(desc(Statistics))%>%kbl(
  caption = "<center><strong>Two-sample K-S Test: cloudy VS cloud-free<center><strong>",
  escape = FALSE,
  format = 'html',
  table.attr = "style='width:45%;'") %>%
  kable_classic(full_width = T, html_font = "Cambria",) 

```
Hence, Author defined variables are the best for the classification task.\

We will likely obtain better classification result using the original author-defined variables: NDAI, CORR, SD than the PCs.\


# Preparation: Data Splitting
Performance might be affected by spatial autocorrelation.

## First Method: 




















