---
title: "proj2"
author: "Kwong Yu Chong, Jiahua Wang"
date: "2022-11-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
require(tidyverse)
require(kableExtra)
require(patchwork)
require(GGally)
require(corrplot)
require(RColorBrewer)
require(tidymodels)
```

# Read Data
```{r}
img1 = read.table("imagem1.txt")
img2 = read.table("imagem2.txt")
img3 = read.table("imagem3.txt")
```

```{r}
col_name = c("Y","X","Label","NDAI","SD","CORR","Rad_Df",
             "Rad_Cf","Rad_Bf","Rad_Af","Rad_An")
colnames(img1) = col_name
colnames(img2) = col_name
colnames(img3) = col_name

img1$image = 1
img2$image = 2
img3$image = 3

full_img = rbind(img1, img2, img3)
```

```{r}
full_img %>% head() 
```

# Summarize the data

###  % pixcel by label
```{r}
tbl1= full_img %>% filter(image==1) %>% group_by(Label) %>% count() %>%
  ungroup() %>% 
  mutate(Image1 = round(n / sum(n)*100,1)) %>% select(Label, Image1)
tbl2= full_img %>% filter(image==2) %>% group_by(Label) %>% count() %>%
  ungroup() %>% 
  mutate(Image2 = round(n / sum(n)*100,1)) %>% select(Label, Image2)
tbl3= full_img %>% filter(image==3) %>% group_by(Label) %>% count() %>%
  ungroup() %>% 
  mutate(Image3 = round(n / sum(n)*100,1)) %>% select(Label, Image3)

tbl1 %>% left_join(tbl2, by='Label') %>% left_join(tbl3, by='Label') %>%
  mutate(Label=ifelse(Label==1, 'Cloudy (1)', ifelse(Label==-1, 'Cloud-free (-1)', 'Unlabeled (0)')))%>%
  arrange(Label)%>%
  kbl(caption = "<center><strong>Pixels By Label (%)<center><strong>",
      escape = FALSE,
      format = 'html',) %>%
  kable_classic(full_width = F, html_font = "Cambria")
```

### Plot Maps
```{r}
full_img %>% filter(image==1) %>% 
  mutate(Label=
           as.factor(ifelse(Label==1, 'Cloudy (1)',
                  ifelse(Label==-1, 'Cloud-free (-1)',
                         'Unlabeled (0)')))) %>% 
  ggplot() +
  geom_raster(aes(x=X, y=Y, fill=Label)) +
  scale_fill_manual(values=c("Cloudy (1)"='white', 
                             'Cloud-free (-1)'='grey',
                             'Unlabeled (0)'='black'))+
  labs(title='Image 1 Label', x='X-coordinate', y='Y-coordinate')+
   theme(plot.title = element_text(hjust = 0.5), legend.position="none")

full_img %>% filter(image==2) %>% 
  mutate(Label=
           as.factor(ifelse(Label==1, 'Cloudy (1)',
                  ifelse(Label==-1, 'Cloud-free (-1)',
                         'Unlabeled (0)')))) %>% 
  ggplot() +
  geom_raster(aes(x=X, y=Y, fill=Label)) +
  scale_fill_manual(values=c("Cloudy (1)"='white', 
                             'Cloud-free (-1)'='grey',
                             'Unlabeled (0)'='black'))+
  labs(title='Image 2 Label', x='X-coordinate', y='Y-coordinate')+
   theme(plot.title = element_text(hjust = 0.5), legend.position="none")

full_img %>% filter(image==3) %>% 
  mutate(Label=
           as.factor(ifelse(Label==1, 'Cloudy (1)',
                  ifelse(Label==-1, 'Cloud-free (-1)',
                         'Unlabeled (0)')))) %>% 
  ggplot() +
  geom_raster(aes(x=X, y=Y, fill=Label)) +
  scale_fill_manual(values=c("Cloudy (1)"='white', 
                             'Cloud-free (-1)'='grey',
                             'Unlabeled (0)'='black'))+
  labs(title='Image 3 Label', x='X-coordinate', y='Y-coordinate')+
   theme(plot.title = element_text(hjust = 0.5),legend.position="none")
```

# EDA
```{r}
col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
full_img %>% select(-c(X,Y,Label, image)) %>% cor() %>%
  corrplot::corrplot(method="color", col=col(200),  
     diag=T,  tl.srt=45, tl.col = 'black', tl.cex = 0.75,mar=c(0,0,2,0),
     type="upper", order="hclust", 
     title='Feature Correlation', 
         addCoef.col = "black",  outline=T,
         )
```

```{r}
tbl_data = full_img %>% select(-image)%>% filter(Label!=0) %>% cor() %>% .[-c(1,2,3),3] %>% round(.,2)

data.frame('value'=tbl_data) %>% arrange(desc(value))%>%kbl(
  caption = "<center><strong>Correlation with Label<center><strong>",
  escape = FALSE,
  format = 'html',
  table.attr = "style='width:25%;'") %>%
  kable_classic(full_width = T, html_font = "Cambria",) 
```

```{r}
plot_data = full_img %>% 
  filter(Label != 0) %>%
  mutate(Label = ifelse(Label==1, 'Cloudy', 'Cloud-free'))

p1 = plot_data%>%
  ggplot() + geom_density(aes(x=CORR, fill=Label), alpha=0.5) 
p2 = plot_data%>%
  ggplot() + geom_density(aes(x=NDAI, fill=Label), alpha=0.5) 
p3 = plot_data%>%
  ggplot() + geom_density(aes(x=log(SD), fill=Label), alpha=0.5) 
p4 = plot_data%>%
  ggplot() + geom_density(aes(x=Rad_Af, fill=Label), alpha=0.5) 

p1 + p2 + p3 + p4 + plot_layout(nrow=2, guides = "collect") + 
  plot_annotation(title = 'Density of Features By Label',
                  theme=theme(plot.title = element_text(hjust = 0.5))) 
```

## Two-sample Kolmogorov–Smirnov statistics
```{r}
full_img_cloudy = full_img %>% select(-image)%>%filter(Label==1)
full_img_not_cloudy = full_img %>% select(-image)%>% filter(Label==-1)

var_names = full_img %>% select(-c(X, Y, Label, image)) %>% colnames() 
KS_stat = matrix(nrow=length(var_names))
rownames(KS_stat) = var_names
colnames(KS_stat) = 'Statistics'

for (var in var_names){
  cloudy_density = full_img_cloudy %>% select(var) %>% as.matrix()%>% density()
  not_cloudy_density = full_img_not_cloudy %>% select(var)%>% as.matrix()%>% density()
  min_x = min(cloudy_density$x , not_cloudy_density$x)
  max_x = max(cloudy_density$x , not_cloudy_density$x)
  cloudy_density = full_img_cloudy %>% select(var) %>% as.matrix()%>% 
    density(from=min_x, to=max_x)
  not_cloudy_density = full_img_not_cloudy %>% select(var)%>% as.matrix()%>%
    density(from=min_x, to=max_x)
  
  x_diff = not_cloudy_density$x %>% diff() %>% .[1]
  KS = tibble(x=cloudy_density$x, cloudy=cloudy_density$y, not_cloudy= not_cloudy_density$y)%>%
    mutate(cloudy = cumsum(cloudy *x_diff),
           not_cloudy= cumsum(not_cloudy *x_diff))%>%
    mutate(diff = abs(cloudy - not_cloudy))%>%
    select(diff)%>%max()
  KS_stat[var,1] = KS
}
KS_stat %>% round(.,2) %>% data.frame() %>% arrange(desc(Statistics))%>%kbl(
  caption = "<center><strong>Two-sample K-S Test: cloudy VS cloud-free<center><strong>",
  escape = FALSE,
  format = 'html',
  table.attr = "style='width:45%;'") %>%
  kable_classic(full_width = T, html_font = "Cambria",) 
```

## PCA
Removed unlabeled data
```{r}
pca_data = full_img %>%select(-c(X,Y,Label, image))
pca = prcomp(pca_data, scale.  =T)
explained_var = (pca$sdev^2) %>% cumsum()
explained_var = round(explained_var / explained_var[8]*100,1)

ggplot()+
  geom_point(aes(x=seq(1,8), y=pca$sdev^2))+
  geom_line(aes(x=seq(1,8), y=pca$sdev^2))+
  labs(x='PC', y='Variance', title='Screeplot')
```

```{r}
pca

```
*PC1 is mostly radiance variables. Look at principal direction.


```{r}
full_pca = tibble(data.frame(pca$x))
full_pca$X = full_img$X
full_pca$Y = full_img$Y
full_pca$Label = ifelse(full_img$Label==1,'Cloudy',ifelse(full_img$Label==-1,'Cloud-free',
                                                               'Unlabeled'))
full_pca$image = full_img$image


p1 = full_pca%>% filter(Label!='Unlabeled')%>%ggplot()+
  geom_density(aes(x=PC1, fill=Label), alpha=0.3)+
  labs(title=paste0('PC1(',explained_var[1],'%)'), y='Density', x='')+
  theme(plot.title = element_text(hjust = 0.5))



p2 = full_pca%>% filter(Label!='Unlabeled')%>%ggplot()+
  geom_density(aes(x=PC2, fill=Label), alpha=0.3)+
  labs(title =paste0('PC2 (',explained_var[2]-explained_var[1],'%)'), y='Density', x='')+
  theme(plot.title = element_text(hjust = 0.5))


p3=full_pca %>% filter(image==1) %>% 
  ggplot() +
  geom_raster(aes(x=X, y=Y, fill=PC1)) +
  #labs(title='Image 1: PC1', x='X-coordinate', y='Y-coordinate')+
   theme(plot.title = element_text(hjust = 0.5), legend.position="none")

p4=full_pca %>% filter(image==1) %>% 
  ggplot() +
  geom_raster(aes(x=X, y=Y, fill=PC2)) +
  #labs(title='Image 1: PC2', x='X-coordinate', y='Y-coordinate')+
   theme(plot.title = element_text(hjust = 0.5), legend.position="none")




p1 + p2 + p3+p4+plot_layout(nrow=2, guides = "collect")
```
```{r}
p1=full_pca %>% filter(image==1) %>% 
  ggplot() +
  geom_raster(aes(x=X, y=Y, fill=PC2)) +
  labs(title='Image 1: PC1', x='', y='')+
   theme(plot.title = element_text(hjust = 0.5), legend.position="none",
         axis.text        = element_blank(),
        axis.ticks       = element_blank(),
        axis.title       = element_blank(),
        panel.background = element_blank())


p2=full_pca %>% filter(image==1) %>% 
  ggplot() +
  geom_raster(aes(x=X, y=Y, fill=PC2)) +
  labs(title='Image 1: PC2', x='', y='')+
   theme(plot.title = element_text(hjust = 0.5), legend.position="none",
         axis.text        = element_blank(),
        axis.ticks       = element_blank(),
        axis.title       = element_blank(),
        panel.background = element_blank())
p3 = full_img %>% filter(image==1) %>% 
  ggplot() +
  geom_raster(aes(x=X, y=Y, fill=NDAI)) +
  labs(title='Image 1: NDAI', x='', y='')+
   theme(plot.title = element_text(hjust = 0.5), legend.position="none",
         axis.text        = element_blank(),
        axis.ticks       = element_blank(),
        axis.title       = element_blank(),
        panel.background = element_blank())
p4 = full_img %>% filter(image==1) %>% 
  ggplot() +
  geom_raster(aes(x=X, y=Y, fill=SD)) +
  labs(title='Image 1: SD', x='', y='')+
   theme(plot.title = element_text(hjust = 0.5), legend.position="none",
         axis.text        = element_blank(),
        axis.ticks       = element_blank(),
        axis.title       = element_blank(),
        panel.background = element_blank())

p1+p2+p3+p4 + plot_layout(nrow=2)
```

```{r}
full_pca %>% select(-image)%>% filter(Label!="Unlabeled")%>% 
  mutate(Label=ifelse(Label=='Cloudy', 1, -1)) %>%
  cor() %>% .[1:8,11] %>% round(.,2) %>% 
  data.frame('value'=.) %>% arrange(desc(value))%>%kbl(
  caption = "<center><strong>Correlation with Label<center><strong>",
  escape = FALSE,
  format = 'html',
  table.attr = "style='width:25%;'") %>%
  kable_classic(full_width = T, html_font = "Cambria",) 
```

```{r}
full_PC = data.frame(pca$x) %>% as.tibble() %>% mutate(Label=full_img$Label)
full_PC_cloudy = full_PC %>% filter(Label==1)
full_PC_not_cloudy = full_PC %>% filter(Label==-1)

var_names = full_PC %>% select(-Label)%>% colnames() 
KS_stat = matrix(nrow=length(var_names))
rownames(KS_stat) = var_names
colnames(KS_stat) = 'Statistics'

for (var in var_names){
  cloudy_density = full_PC_cloudy %>% select(var) %>% as.matrix()%>% density()
  not_cloudy_density = full_PC_not_cloudy %>% select(var)%>% as.matrix()%>% density()
  min_x = min(cloudy_density$x , not_cloudy_density$x)
  max_x = max(cloudy_density$x , not_cloudy_density$x)
  cloudy_density = full_PC_cloudy %>% select(var) %>% as.matrix()%>% 
    density(from=min_x, to=max_x)
  not_cloudy_density = full_PC_not_cloudy %>% select(var)%>% as.matrix()%>%
    density(from=min_x, to=max_x)
  
  x_diff = not_cloudy_density$x %>% diff() %>% .[1]
  KS = tibble(x=cloudy_density$x, cloudy=cloudy_density$y, not_cloudy= not_cloudy_density$y)%>%
    mutate(cloudy = cumsum(cloudy *x_diff),
           not_cloudy= cumsum(not_cloudy *x_diff))%>%
    mutate(diff = abs(cloudy - not_cloudy))%>%
    select(diff)%>%max()
  KS_stat[var,1] = KS
}
KS_stat %>% round(.,2) %>% data.frame() %>% arrange(desc(Statistics))%>%kbl(
  caption = "<center><strong>Two-sample K-S Test: cloudy VS cloud-free<center><strong>",
  escape = FALSE,
  format = 'html',
  table.attr = "style='width:45%;'") %>%
  kable_classic(full_width = T, html_font = "Cambria",) 

```
Hence, Author defined variables are the best for the classification task.\

We will likely obtain better classification result using the original author-defined variables: NDAI, CORR, SD than the PCs.\


# Preparation: Data Splitting
Performance might be affected by spatial autocorrelation.

## First Method: Blocking
```{r}
block_split = function(data, K){
  N =data %>% dim() %>% .[1]
  fold_n = round(N/ K)
  remainder = N - (K-1) * fold_n
  block_id = rep(1:(K-1), each= fold_n)
  block_id = c(block_id, rep(K, remainder))
  data['idx'] = 1:(dim(data)[1])
  data = data %>% arrange(X, Y)
  data['block']= block_id
  
  idx = vector(mode = 'list',length = K)
  for (i in 1:K){
    idx[[i]] = data %>% filter(block==i) %>% pull(idx) 
  }
  return(idx)
}

plot_block = function(data, K, img){
  res = block_split(data, K)
  
  data['block'] = NA
  for (i in 1:K){
    data[res[[i]],'block'] = i
  }
  my.cols = brewer.pal(K, "Paired")
  my.cols = c(my.cols, "white")
  data %>% filter(image==img) %>% 
    mutate(color = ifelse(Label!=0, block, 'Cloud-free'))%>%
    ggplot() +
    geom_raster(aes(x=X, y=Y, fill=color)) +
    labs(title='Image 1: blocks', x='X-coordinate', y='Y-coordinate')+
    scale_fill_manual(values=my.cols)+
    theme(plot.title = element_text(hjust = 0.5), legend.position="none")
}

set.seed(1234567)
plot_block(full_img, 3, 1)

```


## Second Method: KMeans
```{r}
Kmeans_block_split = function(data, K){
  idx = vector(mode = 'list',length = K)
  res = kmeans(data %>%  select(Y,X) ,K)
  for (i in 1:K){
    idx[[i]] = which(res$cluster == i)
  }
  return (idx)
}

plot_kmean_block = function(data, K, img){
  data_copy = rbind(data)
  res = Kmeans_block_split(data, K)
  data_copy['block'] = NA
  for (i in 1:K){
    data_copy[res[[i]],'block'] = i
  }
  my.cols = brewer.pal(K, "Paired")
  my.cols = c(my.cols, "white")
  data_copy %>% filter(image==img) %>% 
    mutate(color = ifelse(Label!=0, block, 'Cloud-free'))%>%
    ggplot() +
    geom_raster(aes(x=X, y=Y, fill=color)) +
    labs(title='Image 1: K-means blocks', x='X-coordinate', y='Y-coordinate')+
    scale_fill_manual(values=my.cols)+
    theme(plot.title = element_text(hjust = 0.5), legend.position="none")
}
set.seed(12345)
plot_kmean_block(full_img , 3, 1)
```
## Split data into Train / Validation / Test
```{r}
set.seed(666)

create_kmeans_split = function(data){
  idx = Kmeans_block_split(data,3)
  
  for (i in 1:3){
    data[idx[[i]],'block']=i
  }
  
  unique_blocks = data %>% distinct(image, block)
  block_n = unique_blocks %>% dim() %>% .[1]
  val_test_samples = sample(1:block_n, 2)
  
  val_img_block = unique_blocks[val_test_samples[1], ]
  test_img_block = unique_blocks[val_test_samples[2], ]
  
  val = data %>% filter(image==val_img_block$image & block==val_img_block$block)
  test = data %>% filter(image==test_img_block$image & block==test_img_block$block)
  train = data %>% setdiff(val) %>% setdiff(test)
  
  
  return(list(val = val,
              test = test,
              train = train))
  
}

create_block_split = function(data){
  idx = block_split(data,3)
  
  for (i in 1:3){
    data[idx[[i]],'block']=i
  }
  
  unique_blocks = data %>% distinct(image, block)
  block_n = unique_blocks %>% dim() %>% .[1]
  val_test_samples = sample(1:block_n, 2)
  
  val_img_block = unique_blocks[val_test_samples[1], ]
  test_img_block = unique_blocks[val_test_samples[2], ]
  
  val = data %>% filter(image==val_img_block$image & block==val_img_block$block)
  test = data %>% filter(image==test_img_block$image & block==test_img_block$block)
  train = data %>% setdiff(val) %>% setdiff(test)

  return(list(val = val,
              test = test,
              train = train))
  
}


kmeans_split = create_kmeans_split(full_img %>% filter(Label!=0))
bk_split = create_block_split(full_img%>% filter(Label!=0))
```

```{r}
N = full_img %>% filter(Label!=0) %>% dim() %>% .[1]
k_train_pct = (kmeans_split$train %>%dim() %>% .[1]) / N
k_val_pct = (kmeans_split$val %>%dim() %>% .[1]) / N
k_test_pct = (kmeans_split$test %>%dim() %>% .[1]) / N


block_train_pct = (bk_split$train %>%dim() %>% .[1]) / N
block_val_pct = (bk_split$val %>%dim() %>% .[1]) / N
block_test_pct = (bk_split$test %>%dim() %>% .[1]) / N

split_ratio = matrix(nrow = 3, ncol = 2)

split_ratio[1,1] = k_train_pct
split_ratio[2,1] = k_test_pct
split_ratio[3,1] = k_val_pct
split_ratio[1,2] = block_train_pct
split_ratio[2,2] = block_test_pct
split_ratio[3,2] = block_val_pct

tbl_data = data.frame(split_ratio * 100 ) %>% round(.,1) 
colnames(tbl_data) = c('K means', 'Block split')
rownames(tbl_data) = c('Train', 'Test', 'Validation')
tbl_data %>%kbl(
  caption = "<center><strong>Data Split Ratio Excluding Unlabeled Points<center><strong>",
  escape = FALSE,
  format = 'html',
  table.attr = "style='width:45%;'") %>%
  kable_classic(full_width = T, html_font = "Cambria",) 

```

## Baseline Accuracy
```{r}
kmeans_valid_accuracy <- mean((kmeans_split$val  %>% pull(Label) )== -1)
kmeans_test_accuracy <- mean((kmeans_split$test %>% pull(Label) )== -1)

block_valid_accuracy <- mean((bk_split$val %>% pull(Label) )== -1)
block_test_accuracy <- mean((bk_split$test %>% pull(Label) )== -1)

kmeans_test_accuracy
kmeans_valid_accuracy

block_test_accuracy
block_valid_accuracy

```
Such a trivial classifier will have high average accuracy if the image itself is mainly cloud-free.

## First order Importance

### K means block split: Logistic Regression
```{r}
variable_names = full_img %>% select(-c(X, Y, Label, image)) %>% colnames()
lr_model = logistic_reg() %>% set_engine('glm')

Ktrain_data = kmeans_split$train %>% filter(Label!=0) %>% mutate(Label = as.factor(ifelse(Label==1, 'Cloudy', 'Cloudfree')))
Kvalid_data = kmeans_split$val %>% filter(Label!=0) %>% mutate(Label = as.factor(ifelse(Label==1, 'Cloudy', 'Cloudfree')))

Blocktrain_data = bk_split$train %>% filter(Label!=0) %>% mutate(Label = as.factor(ifelse(Label==1, 'Cloudy', 'Cloudfree')))
Blockvalid_data = bk_split$val %>% filter(Label!=0) %>% mutate(Label = as.factor(ifelse(Label==1, 'Cloudy', 'Cloudfree')))

res = matrix(NA, nrow=length(variable_names), ncol = 3)
rownames(res) = variable_names
colnames(res) = c('AUC: K-means split', 'AUC: Block split', 'AUC: Avg')
  
coeffs = matrix(NA, nrow=length(variable_names), ncol = 3)
rownames(coeffs) = variable_names
colnames(coeffs) = c('Beta: K-means split', 'Beta: Block split', 'Beta: Avg')

for (var in variable_names){
  lr_recipe = recipe(as.formula(paste0('Label ~ ', var)), data = train_data) %>%  step_normalize(all_numeric_predictors())
  lr_work = workflow() %>% add_model(lr_model) %>% add_recipe(lr_recipe)
  lr_fit = lr_work %>% fit(data = Ktrain_data)
  lr_perf = lr_fit %>% augment(new_data = Kvalid_data) %>%
    select(Label, starts_with('.pred')) 
  res[var,1] = lr_perf %>% roc_auc(Label, .pred_Cloudfree) %>% pull(.estimate)
  coeffs[var,1] = broom::tidy(lr_fit) %>% pull(estimate)%>%.[2]
  
  lr_fit = lr_work %>% fit(data = Ktrain_data)
  lr_perf = lr_fit %>% augment(new_data = Blockvalid_data) %>%
    select(Label, starts_with('.pred')) 
  res[var,2] = lr_perf %>% roc_auc(Label, .pred_Cloudfree) %>% pull(.estimate)
  coeffs[var,2] = broom::tidy(lr_fit) %>% pull(estimate)%>%.[2]
  
  res[var,3] = res[var,1]/2 + res[var,2]/2
  coeffs[var,3] = coeffs[var,1]/2 + coeffs[var,2]/2
}

(res*100) %>% round(.,1) %>% as.data.frame() %>% arrange(desc('AUC: Avg')) %>% kbl(
  caption = "<center><strong>Logistic Regression Performance<center><strong>",
  escape = FALSE,
  format = 'html',
  table.attr = "style='width:50%;'") %>%
  kable_classic(full_width = T, html_font = "Cambria",) 
```

```{r}
coeffs %>% round(.,2) %>% as.data.frame() %>% arrange(desc('Beta: Avg')) %>% kbl(
  caption = "<center><strong>Logistic Regression Coefficient Estimate<center><strong>",
  escape = FALSE,
  format = 'html',
  table.attr = "style='width:50%;'") %>%
  kable_classic(full_width = T, html_font = "Cambria",) 
```
AUC is highly variable on randomization seed, which changes the Train-Validation split. However, NDAI still outperforms under most of the cases.
















